---
title: "class notes11.5"
format: html
author: Mio Takata
embed-resources: true
---

index:

```{r}
#| label: setup
#| message: false
library(tidyverse)
```

A college says that their into stats students are 50% freshman, 30% sophmore, 10% junior, 10%senior. (Population)

- Null hypothesis: those proportions are correct 
- Alternative hypothesis: thy are not correct

Let's get some data in the sample of n = 65 students

fr:28 So:24 Jr:9 Sr:4(0.062)

we would like to know is our data close enough 

obs: 28, 24, 9, 4 (observed values)
exp: 32.5, 19.5, 6.5, 6.5 (expected values under the null hypothesis)

we compute **a chi square test statistics**

- **A goodness-of-fit** test compares a single categorical variable's observed frequencies to expected frequencies from a known distribution, 

- **A test of independence** assesses whether two categorical variables are associated or independent

$\chi^2$  = ($\Sigma(Obs-Exp)^2/Exp$) we devide variance by expected value 
this has $\chi^2$ distribution with k-1 degrees of freedom when k = # of categories

## A goodness of fit test
In R:
```{r}
o <- c(28, 24, 9, 4)
p_null <- c(.5, .3, .1, .1)
chisq.test(o, p = p_null) #it needs to know the data and null hypothesis
```

$\chi^2$  = 3.58
df = 3
p = 0.31 <- big

If $H_0$ is true, data like this occurs 31% of the time.
lower p-values are **evidence against** the null hypothesis. This data is consistent with the null.

**bar graph is needed for Chi square test**

**Do not say** that the null hypothesis is true

