---
title: "Takata_HW11.5"
format: html
author: Mio Takata
embed-resources: true
---

## Problem1
```{r}
#| label: setup
#| message: false
library(tidyverse)
carseats <- ISLR2::Carseats
glimpse(carseats)

ggplot(carseats, aes(x = Price))+
  geom_histogram()+
  geom_vline(xintercept  = 110)+
  theme_minimal()
```
It looks like most of the data points are more than 110
```{r}
t.test(carseats$Price, mu = 110)
```

Null hypothesis: mu = 110($H_0:\mu = 110$)

Alternative hypothesis: mu is not 110($H_a: \mu \ne 110$)

The p-value measure the probability of getting an x-bar this far away from the null hypothesis $\mu$

This small p-value(1.429e-06) provide evidence against the company’s claim that the average price of their own seats is $110(null-hypothesis)

We conclude the mean price is not 110

## Problem2
Compute a level 90% confidence interval for the mean competitor price in the population from which this sample is drawn. What is the sample mean? What is the margin of error?

```{r}
mean(carseats$Price,
     na.rm = TRUE)
```
95 percent confidence interval:
113.4677 118.1223

sample mean:
115.795

margin of error = 115.795 - 113.4677 = 2.327

## Problem3
```{r}
table1<-table(penguins$species, penguins$island) %>% 
  addmargins()
table1
```

## Problem4
[to my github](https://github.com/myosin-t/Math250-github-sandbox/tree/main)




